---
title: "Provocations on Opportunities & Risks:\nGlobal Poverty & Demography"
author: "Bryce Quillin"
date: "8/7/2019"
output:
  beamer_presentation: 
    keep_tex: yes
  slidy_presentation: default
  ioslides_presentation: default
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```


# Overview of Provocations - Opportunities
**New near-term opportunities and risks are being created by massive structural changes in the size, nature and geographical distribution of global demand for goods and services.**

**

* On the opportunities side, extreme global poverty will likely be eradicated in this generation. 
* Data are slow to materialize, but the latest figures from a few years ago show global poverty rates falling below 10% for the first time ever. 
* This is down from almost 30% in 2000, 35% in 1990 and over 40% in 1980. 
* Less progress has been made in addressing non-income dimensions of poverty, but this alone represents the greatest increase in aggregate global demand and purchasing power in history will continue to grow as income above subsistence needs continues to rise.

# Overview of Provocations - Risks
* Yet, on the risks side, changing demographic trends will fundamentally change the nature of demand in hard to predict ways that are not yet well appreciated. Demographic change is not an issue for the distant future and it is impacting a broad range of countries. Labor markets in Brazil, for example are already feeling the effects.
* The world population is growing more slowly and aging at an unprecedented speed. After increasing for five decades, the population of people ages 15-64 (the typical working age) peaked 2012 and is falling. 
* These global trends--slower population growth and aging--are hitting the enginges of growth in the global economy, not just advanced economies but emerging markets across Asia, Europe and Latin America. 
* Many EM will grow old before they grow rich. Asset prices will adjust as retirees draw down pensions, with implications for the costs and availability of global capital. The nature of demand for goods and services will need to adjust and the demand for non-traded services will grow faster than traded products. 

# Global demand for goods and services is rising as more households earn above subsistence level incomes
```{r poverty, echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(ggplot2)
library(ggthemes)

poverty <- read.csv("poverty.csv", sep = ",", header = T)

ggplot(data = poverty) +
  aes(x = year, y = poverty) +
  geom_line(color = '#a69d75') +
  labs(title = 'Global Extreme Poverty Rate (%)',
       caption = 'Source: Pfizer Global Economics with data from World Bank Group.',
       subtitle = 'Share of global population living below $1.90 a day (2011 PPP)') +
  theme_economist() +
  xlab("") + ylab("") +
  geom_text(aes(label = poverty, vjust = -0.5)) +
  theme(plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))

```

# The composition of demand will change as countries across the world age.
```{r wap, echo=FALSE,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# packages
library(tidyverse)
library(rvest)
library(magrittr)
library(ggmap)
library(stringr)
library(maps)
library(ggthemes)

# read data
wap <- read.csv("wap.csv", sep = ",", header = T)

# get world map
map.world <- map_data("world")

# rename USA & UK to match map
wap$country <- recode(wap$country,
                      'United States of America' = 'USA',
                      'United Kingdom' = 'UK',
                      'Russian Federation' = 'Russia',
                      'Iran (Islamic Republic of)' = 'Iran',
                      'Venezuela (Bolivarian Republic of)' = 'Venezuela',
                      'Bolivia (Plurinational State of)' = 'Bolivia',
                      'Viet Nam' = 'Vietnam',
                      'Syrian Arab Republic' = 'Syria',
                      'United Republic of Tanzania' = 'Tanzania',
                      'Congo' = 'Republic of Congo',
                      'Czechia' = 'Czech Republic',
                      'Republic of Moldova' = 'Moldova',
                      'North Macedonia' = 'Macedonia',
                      'Republic of Korea' = 'South Korea',
                        )

# subset out Anarctica
map.world.subset <- subset(map.world, !(region == 'Antarctica'))

# join wap to map
map.world.joined <- left_join(map.world.subset, wap, by = c('region' = 'country'))

# plot map
wap_map <- ggplot() +
  geom_polygon(data = map.world.joined, aes(x = long, y = lat, group = group, fill = change)) +
  xlab("") + ylab("") +
  #labs(caption = "Source: UN Population Division.") + 
  theme_economist() +
  theme(plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5)) +
  theme(plot.caption = element_text(hjust = 0.0)) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_blank() 
  )

ggsave("wap2.png", wap_map, dpi =1000)

```

```{r Cartogram Map, echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# loading libraries
library(devtools)
install_github('omegahat/Rcartogram')
install_github('chrisbrunsdon/getcartr', subdir='getcartr')
library(ggplot2)
library(maptools)
library(rgeos)
library(Rcartogram)
library(getcartr)
library(fftw)
library(rgdal)
library(sp)
library(rgeos)

# load world map & working-age share population data
world <- readShapePoly('TM_WORLD_BORDERS-0.3.shp') # These data were obtained from: http://thematicmapping.org/downloads/world_borders.php
wap <- read.csv("wap.csv", sep = ",", header = T)
# wap$change <- scale(wap$change)

# join the map with the working-age population data
matched.indices <- match(world@data[, "UN"], wap[, "code"])
world@data <- data.frame(world@data, wap[matched.indices, ])
world@data <- na.omit(world@data) # remove missing values as map has some countries not included in pop data

#compute cartogram transformation of each country using change in working-age pop with degree of Gaussian blur = 0.5
world@proj4string <- CRS(as.character(NA))
world.carto <- quick.carto(world, world@data$change, blur = 0.5)




```

```{r Cartogram Map 2, echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(sf)             # spatial data classes
library(rnaturalearth)  # world map data
library(readxl)         # reading excel files
library(dplyr)          # data manipulation
library(tidyr)          # data manipulation
library(purrr)          # data manipulation
library(cartogram)      # cartograms creation
library(tmap)           # maps creation

# load data
world_map = ne_countries(returnclass = "sf")


if(!dir.exists("data")) dir.create("data")
download.file("http://gapm.io/dl_pop", destfile = "data/pop1800_2100.xlsx")
world_pop = read_xlsx("data/pop1800_2100.xlsx", sheet = 7)

# cleaning
world_map = world_map %>% 
  select(sovereignt) %>% 
  filter(sovereignt != "Antarctica") %>% 
  st_transform(world_map, crs = "+proj=robin")

world_pop = world_pop %>% 
  mutate(sovereignt = name) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Tanzania", "United Republic of Tanzania")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "United States", "United States of America")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Congo, Dem. Rep.", "Democratic Republic of the Congo")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Bahamas", "The Bahamas")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Serbia", "Republic of Serbia")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Macedonia, FYR", "Macedonia")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Slovak Republic", "Slovakia")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Czech Republic", "Czechia")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Congo, Rep.", "Republic of the Congo")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Kyrgyz Republic", "Kyrgyzstan")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Lao", "Laos")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Swaziland", "eSwatini")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Cote d'Ivoire", "Ivory Coast")) %>% 
  mutate(sovereignt = replace(sovereignt, sovereignt == "Timor-Leste", "East Timor"))

# prepare
world_data = left_join(world_map, world_pop, by = "sovereignt") %>% 
  na.omit() %>% 
  select(-geo, -name, -indicator) %>% 
  gather(key = "year", value = "population", `1800.0`:`2100.0`) %>% 
  mutate(year = as.integer(year))

world_data2 <- left_join(world_map, wap2, by = "sovereignt") %>% 
  na.omit()

world_data = world_data %>% 
  group_by(year) %>% 
  mutate(total_pop = sum(as.numeric(population), na.rm = TRUE)) %>% 
  mutate(title = paste0("Year: ", year, "\nTotal population (mln): ", round(total_pop/1e6, 2)))

# subset
world_data = world_data %>% 
  filter(year %in% seq(1800, 2100, by = 25))

# create map
world_data_2100 = world_data %>%
  filter(year == 2100)
world_carto1 = cartogram_cont(world_data, "population", maxSizeError = 1.5)
plot(world_carto1["population"])


world_carto2 = cartogram_cont(world_data2, "change", maxSizeError = 1.5)
plot(world_carto1["population"])



############################
wap <- read.csv("wap.csv", sep = ",", header = T)

# cleaning
world_map = world_map %>% 
  select(sovereignt) %>% 
  filter(sovereignt != "Antarctica") %>% 
  st_transform(world_map, crs = "+proj=robin")

# join
map.world.joined <- left_join(world_map, wap, by = c('sovereignt' = 'country')) %>% 
  na.omit %>% 
  select(-code, -Parent.code) 

# create cartogram
world_carto <- cartogram_cont(map.world.joined, "change", maxSizeError = 1.5)

```



```{r}
contributions <- read.csv("contributions.csv", sep = ",", header = T)

contributions_ae <- ggplot(contributions, aes(x = years, y = contributions)) + geom_col(aes(fill = variable)) +
  theme_economist() +
  ylab("Annual Potential GDP growth (%)") + xlab("") +
  labs(title = "Advanced Economies") +
  theme(plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5)) +
  theme(plot.caption = element_text(hjust = 0.0)) +
  theme(legend.title = element_blank()) +
  #scale_fill_brewer(palette = 3) +
  theme(legend.text = element_text(size = 9)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggsave("contributions_ae.png", contributions_ae, dpi = 2000)
 
```


```{r Contributions EM}
library(ggplot2)
library(ggthemes)

contributions_em <- read.csv("contributions_em.csv", sep = ",", header = T)


contributions_em <- ggplot(contributions_em, aes(x = years, y = contributions)) + geom_col(aes(fill = variable)) +
  theme_economist() +
  xlab("") + ylab("") +
  labs(title = "Emerging Market Economies") +
  theme(plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5)) +
  theme(plot.caption = element_text(hjust = 0.0)) +
  theme(legend.title = element_blank()) +
  theme(legend.text = element_text(size = 9)) +
  #scale_fill_brewer(palette = 3) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggsave("contributions_em.png", dpi = 2000)

```

```{r}
library(ggpubr)

panel <- ggarrange(contributions_ae, contributions_em, ncol = 2, nrow = 1, common.legend = T, legend = "bottom") 
ggsave("contributions_panel.png", panel, dpi = 2000)

 
```

